Smart Spreading Factor Assignment for LoRaWANs

### 2019-01-23 paper.tex
hoca name
grammer

\author{\IEEEauthorblockN{Tugrul Yatagan and Sema F. Oktug}
\IEEEauthorblockA{Department of Computer Engineering,
Istanbul Technical University\\
Istanbul, Turkey\\
Email: \{yatagan, oktug\}@itu.edu.tr}}
\maketitle

@@

\par In a LoRaWAN network, initially, a node is not aware of how far away it is from a GW. However, a node can guess the distance from a GW by observing received signal power of a downlink transmission. If received signal power of a downlink transmission is too high, then node can decrease its next transmission SF to decrease power consumption. This SF assignment method is called lowest possible SF assignment scheme for the rest of the paper. Lowest possible SF assignment scheme is commonly used in LoRaWAN deployments. GW can request from a node to decrease its SF or transmit power.

\par In Figure \ref{fig:collision}, a LoRaWAN network deployed with a single GW is illustrated. Different color rings represent achievable range of different SFs from the GW and different color circles represent selected SF of the nodes. The end devices close to the GW will fall into lowest SF (SF7) area section. The end devices close to the GW will probably select the lowest SF mot the time. Hence the number of collisions will increase as the number of end devices close to the GW increases.

@@

\section{LoRaWAN} \label{LoRaWAN}
\par LoRa has an open standard medium access control (MAC) layer protocol called LoRaWAN is which is designed for large scale LoRa networks considering well known LPWAN challenges and their best practice solutions. LoRaWAN is developed and maintained by LoRa Alliance. LoRa Alliance is an open, non-profit organization dedicated to standardization of LoRaWAN. LoRaWAN is based on pure ALOHA medium access which means that end nodes do not check whether the channel is free or not before transmission, accepting the possibility of a collision. A typical LoRaWAN network consists of following three network entities.

@@

\par In \cite{8267219}, the authors studied the effects of imperfect orthogonality between different LoRa SF transmissions. The authors state that a LoRa transmission can be interfered by a different SF transmission when power of the interfering signal significantly is greater than the reference signal. Their experimental results show that this power difference is around 16 dB such a power difference can be seen when an interfering signal is close to a receiver or the sum of interfering signals' energy can create this power difference.

@@

\par It is possible to generate mass amount of LoRaWAN transmission records for different topologies using our simulator. Data set size is directly proportional to simulation duration. Thus, increasing the simulation duration, improves the prediction accuracy. In real world deployments, NS can keep track of all transmissions and it can create a classifier daily basis, then GWs can request from nodes to use suggested SFs.

@@

\par In Figure \ref{fig:gw_pdr}, PDR plot for various number of GWs is shown. Randomly generated topology radius is set to 3000 meters and the lowest SF assignment scheme is used. Increasing number of GWs, decreases the SFs of nodes and decreases air time. This decreases number of collisions, henc, increases the PDR of network when network radius is constant.

@@

\par In Table \ref{table:prediction_accuracy}, prediction accuracy of smart SVM and smart DTC schemes are presented for various network radii and number of nodes. Prediction accuracy is not directly proportional to network PDR. Correct prediction of an interfered transmission may not increase the PDR but increases the accuracy. Mart DTC gives better PDR results than smart SVM, even overall SVM prediction accuracy is high.

@@

\begin{table}
\centering
\caption{Prediction accuracy of SVM and DTC classifiers.}
\label{table:prediction_accuracy}
\begin{tabular}{|c|c|c|c|c|c|}
\hline
\multicolumn{3}{|c|}{\multirow{2}{*}{}}                        & \multicolumn{3}{c|}{\textbf{Number of Nodes}} \\ \cline{4-6}
\multicolumn{3}{|c|}{}                                         & 100           & 500           & 1000          \\ \hline
\multirow{8}{*}{\textbf{R (m)}} & \multirow{2}{*}{3000}  & SVM & 82.4          & 70.4          & 71.7          \\ \cline{3-6}
                                &                        & DTC & 86.0          & 67.3          & 70.4          \\ \cline{2-6}

@@

\par For space constraints, we omit results like network throughput and total transmit energy consumption. We invite readers to experiment the simulation tool with the parameters they like. \cite{simlorasf}.

\section{Conclusion} \label{Conclusion}
\par In this paper, we present a discrete event simulator which is developed from scratch to study network performance of LoRaWAN and evaluate different SF assignment schemes. Moreover, we show how SF collisions can be avoided, hence, we propose a machine learning based solutions called smart DTC scheme and smart SVM scheme. We present simulation results for the lowest SF assignment scheme and the proposed schemes. Simulation results show that, the proposed smart schemes can increase network performance for LoRaWAN networks, especially, when the nodes are deployed close to GWs.

\par As for future work, transmit power optimization can be included to the proposed smart schemes. In this paper, it is assumed that nodes always use maximum transmit power for uplink transmission, however nodes close to GWs can decrease transmit power to save energy. This will make transmissions more vulnerable to interference thus requires extra care. Moreover, other machine learning methods can be investigated for SF assignment enhancement. Reinforcement learning could be a good candidate. Also, other transmission parameters such as node id and transmission time can be included to the proposed scheme to the improve performance.

>>>>

\author{\IEEEauthorblockN{Tugrul Yatagan and Sema Oktug}
\IEEEauthorblockA{Department of Computer Engineering,
Istanbul Technical University\\
Istanbul, Turkey\\
Email: \{yatagan, oktug\}@itu.edu.tr}}
\maketitle

@@

\par In a LoRaWAN network, initially, a node is not aware of how far away it is from a GW. However, a node can guess the distance from a GW by observing received signal power of a downlink transmission. If received signal power of a downlink transmission is too high, then node can decrease its next transmission SF to decrease power consumption. This SF assignment method is called lowest possible SF assignment scheme for the rest of the paper. Lowest possible SF assignment scheme is commonly used in LoRaWAN deployments. Also, a GW can request from a node to decrease its SF or transmit power.

\par In Figure \ref{fig:collision}, a LoRaWAN network deployed with a single GW is illustrated. Different color rings represent achievable range of different SFs from the GW and different color circles represent selected SF of the nodes. The end devices close to the GW will fall into lowest SF (SF7) area section. The end devices close to the GW will probably select the lowest SF most of the time. Hence the number of collisions will increase as the number of end devices close to the GW increases.

@@

\section{LoRaWAN} \label{LoRaWAN}
\par LoRa has an open standard medium access control (MAC) layer protocol called LoRaWAN which is designed for large scale LoRa networks considering well known LPWAN challenges and their best practice solutions. LoRaWAN is developed and maintained by LoRa Alliance. LoRa Alliance is an open, non-profit organization dedicated to standardization of LoRaWAN. LoRaWAN is based on pure ALOHA medium access which means that end nodes do not check whether the channel is free or not before transmission, accepting the possibility of a collision. A typical LoRaWAN network consists of following three network entities.

@@

\par In \cite{8267219}, the authors studied the effects of imperfect orthogonality between different LoRa SF transmissions. The authors state that a LoRa transmission can be interfered by a different SF transmission when power of the interfering signal is significantly greater than the reference signal. Their experimental results show that this power difference is around 16 dB. Such a power difference can be seen when an interfering signal is close to a receiver or the sum of interfering signals' energy can create this power difference.

@@

\par It is possible to generate mass amount of LoRaWAN transmission logs for different topologies using our simulator. Data set size is directly proportional to simulation duration. Thus, increasing the simulation duration, improves the prediction accuracy. In real world deployments, NS can keep track of  transmissions and it can create a classifier daily basis, then GWs can request from nodes to use suggested SFs.

@@

\par In Figure \ref{fig:gw_pdr}, PDR plot for various number of GWs is shown. Randomly generated topology radius is set to 3000 meters and the lowest SF assignment scheme is used. Increasing number of GWs, decreases the SFs of nodes and decreases air time. This decreases number of collisions, hence, increases the PDR of network when network radius is constant.

@@

\par In Table \ref{table:prediction_accuracy}, prediction accuracy of smart SVM and smart DTC schemes are presented for various network radii and number of nodes. Prediction accuracy is not directly proportional to network PDR. Correct prediction of an interfered transmission may not increase the PDR but increases the accuracy. Smart DTC gives better PDR results than smart SVM, even overall SVM prediction accuracy is high.

@@

\begin{table}
\centering
\caption{Prediction accuracy for SVM and DTC.}
\label{table:prediction_accuracy}
\begin{tabular}{|c|c|c|c|c|c|}
\hline
\multicolumn{3}{|c|}{\multirow{2}{*}{}}                        & \multicolumn{3}{c|}{\textbf{Number of Nodes}} \\ \cline{4-6}
\multicolumn{3}{|c|}{}                                         & 100           & 500           & 1000          \\ \hline
\multirow{8}{*}{\textbf{R (m)}} & \multirow{2}{*}{3000}  & SVM & 82.4          & 70.4          & 71.7          \\ \cline{3-6}
                                &                        & DTC & 86.0          & 67.3          & 70.4          \\ \cline{2-6}

@@

\par For space constraints, we omit results like network throughput and total transmit energy consumption. We invite readers to experiment the simulation tool with the parameters they like \cite{simlorasf}.

\section{Conclusion} \label{Conclusion}
\par In this paper, we present a discrete event simulator which is developed from scratch to study network performance of LoRaWAN and evaluate different SF assignment schemes. Moreover, we show how same SF collisions can be avoided, hence, we propose a machine learning based solutions called smart DTC scheme and smart SVM scheme. We present simulation results for the lowest SF assignment scheme and the proposed schemes. Simulation results show that, the proposed smart schemes can increase network performance for LoRaWAN networks, especially, when the nodes are deployed close to GWs.

\par As for future work, transmit power optimization can be included to the proposed smart schemes. In this paper, it is assumed that nodes always use maximum transmit power for uplink transmission, however nodes close to GWs can decrease transmit power to save energy. This will make transmissions more vulnerable to interference thus requires extra care. Moreover, other machine learning methods can be investigated for SF assignment enhancement. Reinforcement learning could be a good candidate. Also, other transmission parameters such as node id and transmission time can be included to the proposed scheme to the improve prediction performance.
